### 2013-12-19 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Working on DPP ms.
Finishing incorporation of the new figures. Also, updating some of the
new figures to add an extra digit for the exponential distributions of
divergence times.

We find that all four models accurately estimate the posterior probability of
the one-divergence model when applied to their own datasets (i.e., when the
prior is correct; see diagonal of Figures~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).
The \modelUniform and \modelDPP show robustness to prior violations and
perform well when applied to data generated under other models
(Figures~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).
However, both are less accurate and tends to underestimate the probability of the
one-divergence model when applied to the data generated under \modelUshaped
(Figures~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).
The model-choice behavior of the \modelDPP is relatively accurate when applied
to datasets generated under models \modelDPP, \modelUniform, and \modelOld
(Figs.\
\labelcref{figValidationModelChoiceDPPDPP,figValidationModelChoiceUniformDPP,figValidationModelChoiceOldDPP}).
It is less accurate and tends to underestimate the probability of the
one-divergence model when applied to the data generated under \modelUshaped
(Figure \ref{figValidationModelChoiceUshapedDPP}).
The \modelUniform model performs almost identically as \modelDPP in
terms of model-choice accuracy (Figs.\
\labelcref{figValidationModelChoiceDPPUniform,figValidationModelChoiceUniformUniform,figValidationModelChoiceOldUniform,figValidationModelChoiceUshapedUniform}).

The original \msb model \modelOld performs reasonably well when applied
to its own datasets (Figure \ref{figValidationModelChoiceOldOld}), but
it performs poorly when applied to data generated under any of the other
models (Figs.\
\labelcref{figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld}).
In these cases the \modelOld model is strongly biased towards over-estimating
the posterior probability of the one-divergence model, especially when the
divergence models are not distributed under its U-shaped prior (Figs.\
\labelcref{figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld}).
The other model with the U-shaped prior on divergence models, but exponential
priors on nuisance parameters (\modelUshaped), performs similarly to \modelOld
in that it performs well when applied to its own data, but overestimates
the probability of the one-divergence model when applied to data generated
by the other models (Figs.\
\labelcref{figValidationModelChoiceDPPUshaped,figValidationModelChoiceUniformUshaped,figValidationModelChoiceUshapedUshaped,figValidationModelChoiceUshapedUshaped}).
However, the bias toward over-estimating the probability of the one-divergence
model is stronger in the \modelOld model than \modelUshaped.

@@

\subsection*{Validation analyses: Ordered divergence models}
Our results show that the ordered models of divergence, \modelDPPOrdered,
perform similarly to sampling over unordered models (\modelDPP) when the prior
is correct (Figs.\
\labelcref{figValidationAccuracyDPPOrderedDPPOrdered,figValidationModelChoiceDPPOrderedDPPOrdered}).
This suggests that the method is not adversely affected by the large increase
in the number of possible discrete models (from 22 unordered to 4140 ordered
models) when there are eight pairs of populations.

\subsection*{Power analyses: Estimation accuracy}
All of the models we evaluated (Table \ref{tabPriors}) struggle to estimate the
dispersion index of divergence times \divTimeDispersion regardless of which of
the three series of models (Table \ref{tabPowerModels}) the data were generated
under (Figs.\
\labelcref{figPowerAccuracyExpDPP,figPowerAccuracyExpUniform,figPowerAccuracyExpOld,figPowerAccuracyExpUshaped,figPowerAccuracyUniformDPP,figPowerAccuracyUniformUniform,figPowerAccuracyUniformOld,figPowerAccuracyUniformUshaped,figPowerAccuracyOldDPP,figPowerAccuracyOldUniform,figPowerAccuracyOldOld,figPowerAccuracyOldUshaped}).
When the divergence times of the 22 population pairs are randomly drawn from a
series of exponential priors (\powerSeriesExp), the \modelDPP model is the
best estimator of \divTimeDispersion, followed by \modelUniform (figs.\
\labelcref{figPowerAccuracyExpDPP,figPowerAccuracyExpUniform}).
The \modelOld model is strongly biased toward underestimating
\divTimeDispersion, estimating values of zero for most of the replicates across
all the data models of \powerSeriesExp (Figure \ref{figPowerAccuracyExpOld}).
The results of the \modelUshaped model are intermediate between those of
\modelOld and the new models \modelDPP and \modelUniform
(Figure \ref{figPowerAccuracyExpUshaped}).

When the true divergence times are randomly drawn from a series of uniform
priors (\powerSeriesUniform), all of the models poorly estimate
\divTimeDispersion (Figs.\
\labelcref{figPowerAccuracyUniformDPP,figPowerAccuracyUniformUniform,figPowerAccuracyUniformOld,figPowerAccuracyUniformUshaped}).
The \modelDPP and \modelUniform models tend to over-estimate the variance in
divergence times (Figs.\
\labelcref{figPowerAccuracyUniformDPP,figPowerAccuracyUniformUniform}), whereas
the \modelOld model, again, underestimates \divTimeDispersion, estimating
values of zero for most replicates across all the data models of
\powerSeriesUniform (Figure \ref{figPowerAccuracyUniformOld}).
Again, the performance of the \modelUshaped model is intermediate between the
\modelOld and \modelDPP/\modelUniform models (Figure
\ref{figPowerAccuracyUniformUshaped}).
This pattern of results across the four models is very similar to their
behavior when applied to data simulated under the series of models of
\powerSeriesOld (Figs.\
\labelcref{figPowerAccuracyOldDPP,figPowerAccuracyOldUniform,figPowerAccuracyOldOld,figPowerAccuracyOldUshaped}).

\subsection*{Power analyses: Model choice}
There is a large amount of heterogeneity in estimating the number of divergence
events (\divTimeNum) both among the prior models, and across the different
series of data models (Table \ref{tabPowerModels}) (Figs.\
\labelcref{figPowerPsiExpDPP,figPowerPsiExpUniform,figPowerPsiExpOld,figPowerPsiExpUshaped,figPowerPsiUniformDPP,figPowerPsiUniformUniform,figPowerPsiUniformOld,figPowerPsiUniformUshaped,figPowerPsiOldDPP,figPowerPsiOldUniform,figPowerPsiOldOld,figPowerPsiOldUshaped}).
The \modelDPP model tends to infer highly clustered divergence models across
all three series of data models when the divergences are recent, but performs
the best of the four prior models  when divergences are random over a timescale
of about one to two coalescent units (Figs.\
\labelcref{figPowerPsiExpDPP,figPowerPsiUniformDPP,figPowerPsiOldDPP,}).
The \modelUniform model performs the best when divergences are random
over the most recent timescales, but tends to only infer a tight
cluster of \divTimeNum values even when divergences are random
over longer periods (Figs.\
\labelcref{figPowerPsiExpUniform,figPowerPsiUniformUniform,figPowerPsiOldUniform,}).
The \modelOld model performs the worst of the four prior models across
all three series of data models, inferring a single divergence event across
most of the 18,000 simulations (Figs.\
\labelcref{figPowerPsiExpOld,figPowerPsiUniformOld,figPowerPsiOldOld,}).
Even when the simulated data are identically distributed as the \modelOld model
(except for the divergence times) the \modelOld is still strongly biased
towards a single event (Figure \ref{figPowerPsiOldOld}), whereas the \modelDPP
model is the least likely to infer a single event when applied to these data
(Figure \ref{figPowerPsiOldDPP}).
Using exponential priors on nuisance parameters does increase the power
of the \modelUshaped model compared to \modelOld across all three series
of data models, but the U-shaped prior still prevents the model
from performing as well as the \modelDPP and \modelUniform models (Figs.\
\labelcref{figPowerPsiExpUshaped,figPowerPsiUniformUshaped,figPowerPsiOldUshaped,}).

The improved power of the new models are even more pronounced when looking at
estimates of the dispersion index of divergence times (\divTimeDispersion)
across the simulations (Figs.\
\labelcref{figPowerDispersionExpDPP,figPowerDispersionExpUniform,figPowerDispersionExpOld,figPowerDispersionExpUshaped,figPowerDispersionUniformDPP,figPowerDispersionUniformUniform,figPowerDispersionUniformOld,figPowerDispersionUniformUshaped,figPowerDispersionOldDPP,figPowerDispersionOldUniform,figPowerDispersionOldOld,figPowerDispersionOldUshaped}).
The \modelDPP and \modelUniform models perform similarly across all three
series of data models, inferring values of \divTimeDispersion consistent with
one divergence event ($\divTimeDispersion < 0.01)$ in almost none of the
replicates across all the simulations (Figs.\
\labelcref{figPowerDispersionExpDPP,figPowerDispersionExpUniform,figPowerDispersionUniformDPP,figPowerDispersionUniformUniform,figPowerDispersionOldDPP,figPowerDispersionOldUniform}).
The \modelOld model infers values consistent with a single divergence event in
most of the replicates across all the simulations (Figs.\
\labelcref{figPowerDispersionExpOld,figPowerDispersionUniformOld,figPowerDispersionOldOld}).
Using exponential priors on nuisance parameters greatly increases the power
of the \modelUshaped model relative to \modelOld, but it still has
less power than the models with DPP or uniform priors across
divergence models (Figs.\
\labelcref{figPowerDispersionExpUshaped,figPowerDispersionUniformUshaped,figPowerDispersionOldUshaped}).

When looking at the estimated posterior probability of the single-divergence
model across the power analyses, we also see increased power of the new models
(Figs.\
\labelcref{figPowerPsiProbExpDPP,figPowerPsiProbExpUniform,figPowerPsiProbExpOld,figPowerPsiProbExpUshaped,figPowerPsiProbUniformDPP,figPowerPsiProbUniformUniform,figPowerPsiProbUniformOld,figPowerPsiProbUniformUshaped,figPowerPsiProbOldDPP,figPowerPsiProbOldUniform,figPowerPsiProbOldOld,figPowerPsiProbOldUshaped}).
The \modelDPP model estimates low posterior probability of
$\divTimeNum = 1$ across all of the simulations (Figs.\
\labelcref{figPowerPsiProbExpDPP,figPowerPsiProbUniformDPP,figPowerPsiProbOldDPP}).
The \modelDPP does infer strong support for one divergence as gauged by a Bayes
factor of greater than 10.
However, this is not that unexpected given the low prior probability of a
single divergence under this model.
Also, the Bayes factor calculations for the \modelDPP replicates are only
approximate; they are calculated based on the prior mean of the concentration
parameter \concentrationParam.

The \modelUniform model infers very low posterior probabilities for the
one-divergence model across all the simulations (Figs.\
\labelcref{figPowerPsiProbExpUniform,figPowerPsiProbUniformUniform,figPowerPsiProbOldUniform}),
whereas the \modelOld model infers high posterior probabilities of a single
divergence for most replicates across all simulations (Figs.\
\labelcref{figPowerPsiProbExpOld,figPowerPsiProbUniformOld,figPowerPsiProbOldOld}).
Again, we see that the exponential priors on nuisance parameters greatly
increase power and result in lower estimates of the probability of
one divergence when comparing \modelUshaped to \modelOld (Figs.\
\labelcref{figPowerPsiProbExpUshaped,figPowerPsiProbUniformUshaped,figPowerPsiProbOldUshaped}).

Lastly, when we look at the estimated posterior probability of
\divTimeDispersion being consistent with one divergence parameter
($p(\divTimeDispersion < 0.01 | \ssSpace)$), we see the same pattern of model
behavior, with \modelDPP and \modelUniform inferring low probabilities across
all simulations, \modelOld inferring high probabilities, and \modelUshaped
inferring intermediate values (Figs.\
\labelcref{figPowerDispersionProbExpDPP,figPowerDispersionProbExpUniform,figPowerDispersionProbExpOld,figPowerDispersionProbExpUshaped,figPowerDispersionProbUniformDPP,figPowerDispersionProbUniformUniform,figPowerDispersionProbUniformOld,figPowerDispersionProbUniformUshaped,figPowerDispersionProbOldDPP,figPowerDispersionProbOldUniform,figPowerDispersionProbOldOld,figPowerDispersionProbOldUshaped}).

@@

\section*{Discussion}
Our results demonstrate that using alternative priors on nuisance parameters
and divergence models improved the behavior of the \msb model.
In our implementation, model-choice estimation is more accurate and shows
greater robustness to model violations (Figs.\
\labelcref{figValidationModelChoiceDPPDPP,figValidationModelChoiceUniformDPP,figValidationModelChoiceOldDPP,figValidationModelChoiceUshapedDPP,figValidationModelChoiceDPPUniform,figValidationModelChoiceUniformUniform,figValidationModelChoiceOldUniform,figValidationModelChoiceUshapedUniform,figValidationModelChoiceDPPUshaped,figValidationModelChoiceUniformUshaped,figValidationModelChoiceOldUshaped,figValidationModelChoiceUshapedUshaped,figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld,figValidationModelChoiceOldOld}).
The original model is very sensitive to model violations and, when present,
strongly over-estimates the probability of one divergence event shared
across all taxa (Figs.\
\labelcref{figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld}).
When more appropriate priors are used for nuisance parameters, and either
a Dirichlet-process or uniform prior applied across divergence models,
the model is less sensitive to violations (Figs.\
\labelcref{figValidationModelChoiceUniformDPP,figValidationModelChoiceOldDPP,figValidationModelChoiceUshapedDPP,figValidationModelChoiceDPPUniform,figValidationModelChoiceOldUniform,figValidationModelChoiceUshapedUniform}),
and, when they do cause bias, the method is conservative and tends to
under-estimate the probability of biogeographically interesting models of
temporally clustered divergences (Figs.\
\labelcref{figValidationModelChoiceUshapedDPP,figValidationModelChoiceUshapedUniform}).

Furthermore, the modifications to the model also improve the method's power to
detect random variation in divergence times, reducing the bias toward
estimating clustered divergences (Figs.\
\labelcref{figPowerPsiExpDPP,figPowerPsiExpUniform,figPowerPsiExpOld,figPowerPsiExpUshaped,figPowerPsiUniformDPP,figPowerPsiUniformUniform,figPowerPsiUniformOld,figPowerPsiUniformUshaped,figPowerPsiOldDPP,figPowerPsiOldUniform,figPowerPsiOldOld,figPowerPsiOldUshaped,figPowerDispersionExpDPP,figPowerDispersionExpUniform,figPowerDispersionExpOld,figPowerDispersionExpUshaped,figPowerDispersionUniformDPP,figPowerDispersionUniformUniform,figPowerDispersionUniformOld,figPowerDispersionUniformUshaped,figPowerDispersionOldDPP,figPowerDispersionOldUniform,figPowerDispersionOldOld,figPowerDispersionOldUshaped,figPowerPsiProbExpDPP,figPowerPsiProbExpUniform,figPowerPsiProbExpOld,figPowerPsiProbExpUshaped,figPowerPsiProbUniformDPP,figPowerPsiProbUniformUniform,figPowerPsiProbUniformOld,figPowerPsiProbUniformUshaped,figPowerPsiProbOldDPP,figPowerPsiProbOldUniform,figPowerPsiProbOldOld,figPowerPsiProbOldUshaped,figPowerDispersionProbExpDPP,figPowerDispersionProbExpUniform,figPowerDispersionProbExpOld,figPowerDispersionProbExpUshaped,figPowerDispersionProbUniformDPP,figPowerDispersionProbUniformUniform,figPowerDispersionProbUniformOld,figPowerDispersionProbUniformUshaped,figPowerDispersionProbOldDPP,figPowerDispersionProbOldUniform,figPowerDispersionProbOldOld,figPowerDispersionProbOldUshaped}).
Our results are similar to those of \citet{Oaks2012} in that we also find
the \msb model to infer strong support for clustered divergences when
divergences are random over quite broad timescales (Figs.\
\labelcref{figPowerPsiProbExpOld,figPowerPsiProbUniformOld,figPowerPsiProbOldOld,figPowerDispersionProbExpOld,figPowerDispersionProbUniformOld,figPowerDispersionProbOldOld}).
Our results expand on this by showing that this behavior is consistent
across a range conditions underlying the data.
When using the new priors, the method is less prone to spurious inference of
clustered divergence models, and much less likely to incorrectly infer such
models with strong support.

By evaluating a model intermediate between the old and new implementation
(\modelUshaped), we were able to determine the relative effects of the
modifications to model.
Across all of our analyses, the results show that using better priors on
nuisance parameters alone does improve the performance of the method.
The magnitude of the bias toward inferring support for the one-divergence
model when there are model violations is reduced when the exponential
priors are used in place of the uniform priors (Figs.\
\labelcref{figValidationModelChoiceDPPOld,figValidationModelChoiceUniformOld,figValidationModelChoiceUshapedOld,figValidationModelChoiceDPPUshaped,figValidationModelChoiceUniformUshaped,figValidationModelChoiceOldUshaped}).
Furthermore, better priors on nuisance parameters improves the method's
power to detect temporally random divergences.
Throughout our analyses, the intermediate model (\modelUshaped) performs
better than the \msb model, but not as well as the models with alternative
priors on divergence models.
This suggests, as predicted by \citet{Oaks2012}, that the bias of \msb toward
models of temporally clustered divergences is caused by a combination of
(1) uniform priors on nuisance parameters hindering the marginal likelihoods of
models with more \divTime{} parameters and
(2) the U-shaped prior on divergence models giving low prior density to models
with intermediate numbers of divergence parameters.

While our modifications improve the behavior of the model, we urge
caution when using the method and interpreting its results.
The method attempts to approximate the posterior of a very parameter rich model
using relatively little information from the data.
For example, when applied to the dataset of Philippine vertebrates analyzed by
\citet{Oaks2012} the model can have as many as 604--625 free parameters
(depending on \divTimeNum), and samples over 1002 divergence models.
Our modifications allow this model to be simplified to 471--492 free
parameters.
Sampling over hundreds of parameters and more than a thousand models is a
difficult inference problem, especially when only using a small number of
summary statistics calculated from the sequence alignments of each taxon pair.
The population-genetic summary statistics used by our method contain little
information about many of the free parameters in the model.
Thus, we expect the improved method will still be sensitive to priors, and the
power, while improved, may still be low.
Furthermore, ABC methods in general have been shown to be biased
\citep{Robert2011} for model choice.
As recommended by \citet{Oaks2012}, any results from the method should be
accompanied by assessments of prior sensitivity and power.
We encourage investigators to view this method and its predecessor as a means
of exploring their data rather than a rigorous means of evaluating hypotheses.

>>>>

We find that all four models accurately estimate the posterior probability of
the one-divergence model when applied to their own datasets (i.e., when the
prior is correct; see diagonal of Figures~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).
The \modelUniform and \modelDPP show robustness to prior violations and
perform well when applied to data generated under other models
(Figures~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).
However, both are less accurate and tend to underestimate the probability of the
one-divergence model when applied to the data generated under \modelUshaped
(Figures~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).
In contrast, the original is biased toward overestimating the posterior
probability of the one-divergence model when applied to data generated under
the other three models
(Figures~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).
This bias is particularly strong whenever divergence models are not
distributed under its U-shaped prior
(Figure~\ref{figValidationModelChoicePsi}C--D).
The other model with the U-shaped prior on divergence models, but exponential
priors on parameters (\modelUshaped), performs similarly to the \modelOld model
in that it performs well when applied to its own data, but overestimates the
probability of the one-divergence model when applied to data generated by the
other models
(Figures~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).
However, the bias is stronger in the \modelOld model than \modelUshaped.

@@

\subsection*{Validation analyses: Ordered divergence models}
Our results show that the method performs similarily when sampling over ordered
models of divergence
(Figures S\ref{figValidationAccuracyDPPOrderedDPPOrdered} \&
S\ref{figValidationModelChoiceDPPOrderedDPPOrdered}).
This suggests that the method is not adversely affected by the large increase
in the number of possible discrete models (from 22 unordered to 4140 ordered
models) when there are eight pairs of populations.

\subsection*{Power analyses: Estimation accuracy}
All of the models we evaluated (Table \ref{tabPriors}) struggle to estimate the
variance of divergence times \divTimeDispersion regardless of which of the
three series of models (Table \ref{tabPowerModels}) the data were generated
under
(Figures
S\labelcref{figPowerAccuracyOld,figPowerAccuracyUniform,figPowerAccuracyExp}).
The models with the U-shaped prior on divergence models (\modelOld and
\modelUshaped) tend to underestimate the variance in divergence times 
(Plots A--L of Figures
S\labelcref{figPowerAccuracyOld,figPowerAccuracyUniform,figPowerAccuracyExp}).
whereas the models with Uniform or Dirichlet-process priors over divergence
models tend to overestimate variance in divergence times
(Plots M--X of Figures
S\labelcref{figPowerAccuracyOld,figPowerAccuracyUniform,figPowerAccuracyExp}).

When the divergence times of the 22 population pairs are randomly drawn from a
series of exponential priors (\powerSeriesExp), the \modelDPP model is the
best estimator of \divTimeDispersion, followed by \modelUniform
(Figure~\ref{figPowerAccuracyExp}).
The \modelOld model is strongly biased toward underestimating
\divTimeDispersion, estimating values of zero for most of the replicates across
all the data models of \powerSeriesExp (Figure \ref{figPowerAccuracyExp}).
The results of the \modelUshaped model are intermediate between those of
\modelOld and the new models \modelDPP and \modelUniform
(Figure \ref{figPowerAccuracyExp}).

Similarly, when the true divergence times are randomly drawn from a series of
uniform priors (\powerSeriesUniform), the \modelDPP and \modelUniform models
tend to over-estimate the variance in divergence times, whereas the \modelOld
model, underestimates \divTimeDispersion, estimating values of zero for most
replicates across all the data models of \powerSeriesUniform
(Figure~\ref{figPowerAccuracyUniform}).
Again, the performance of the \modelUshaped model is intermediate between the
\modelOld and \modelDPP/\modelUniform models (Figure
\ref{figPowerAccuracyUniform}).
The results are very similar when the four models are applied to the data
simulated under the \powerSeriesOld seires of models
(Figure~\ref{figPowerAccuracyOld}).

\subsection*{Power analyses: Model choice}
Our modifications of the \msb model decrease the method's bias toward clustered
divergences when applied to data generated under random divergence times
(Figure~\ref{figPowerPsiOld4}A--D \& plots A--F of Figures
S\labelcref{figPowerPsiOld,figPowerPsiUniform,figPowerPsiExp}).
The \modelOld model performs the worst of the four models across
all three series of data-generating models, inferring a single divergence event across
most of the 18,000 simulations
(Figure~\ref{figPowerPsiOld4}A--D \& plots A--F of Figures
S\labelcref{figPowerPsiOld,figPowerPsiUniform,figPowerPsiExp}).
Importantly, the \modelOld tends to strongly support these estimates
of one divergence across most of the simulations
(Figure~\ref{figPowerPsiProbOld4}A--D \& plots A--F of Figures
S\labelcref{figPowerPsiProbOld,figPowerPsiUniform,figPowerPsiExp}).
The \modelDPP model also prefers the one-divergence model 
when divergences are random over narrow windows of time, but
performs much better when divergences are random over a timescale
of 1--2 coalescent units
(Figure~\ref{figPowerPsiOld4}M--P \& plots S--X of Figures
S\labelcref{figPowerPsiOld,figPowerPsiUniform,figPowerPsiExp}).
However, even when \modelDPP infers the one-divergence model over narrow
timescales, the posterior probability support is always low
(Figure~\ref{figPowerPsiProbOld4}M--P \& plots S--X of Figures
S\labelcref{figPowerPsiProbOld,figPowerPsiUniform,figPowerPsiExp}).
% Even when the simulated data are identically distributed as the \modelOld model
% (except for the divergence times) the \modelOld is still strongly biased
% towards a single event (Figure \ref{figPowerPsiOldOld}), whereas the \modelDPP
% model is the least likely to infer a single event when applied to these data
% (Figure \ref{figPowerPsiOldDPP}).
The \modelUniform model never infers the one-divergence model in any of the
simulation replicates but still tends to infer relatively few (4--6) divergence
events when divergences are random over longer periods 
(Figure~\ref{figPowerPsiOld4} I--L \& plots M--R of Figures
S\labelcref{figPowerPsiOld,figPowerPsiUniform,figPowerPsiExp}).
Using exponential priors on divergence-time and theta parameters does increase
the power of the \modelUshaped model compared to \modelOld across all three
series of data models, but the U-shaped prior still prevents the model from
performing as well as the \modelDPP and \modelUniform models 
(Figure~\ref{figPowerPsiOld4} \& 
S\labelcref{figPowerPsiOld,figPowerPsiUniform,figPowerPsiExp}).

The improved power of the new models are even more pronounced when looking at
estimates of the variance of divergence times (\divTimeDispersion) across the
simulations
(Figure~\ref{figPowerOmegaOld4} \& 
S\labelcref{figPowerOmegaOld,figPowerOmegaUniform,figPowerOmegaExp}).
The performance among the models is so different, that the histograms of
\divTimeDispersion estimates cannot be plotted along a shared x-axis.
The \modelDPP and \modelUniform models perform similarly across all three
series of data models, inferring values of \divTimeDispersion consistent with
one divergence event ($\divTimeDispersion < 0.01)$ in almost none of the
replicates across all the simulations.
In contrast, the \modelOld model infers values consistent with a single divergence event in
most of the replicates across all the simulations.
Using exponential priors on divergence-time and theta parameters greatly
increases the power of the \modelUshaped model to detect variation in
divergence times relative to \modelOld, but it still has less power than the
models with Dirichlet-process or uniform priors across divergence models 
(Figure~\ref{figPowerOmegaOld4} \& 
S\labelcref{figPowerOmegaOld,figPowerOmegaUniform,figPowerOmegaExp}).

As mentioned above, we also see increased power of the new models when
looking at the estimated posterior probability of the one-divergence model
across the power analyses 
(Figure~\ref{figPowerPsiProbOld4} \& 
S\labelcref{figPowerPsiProbOld,figPowerPsiUniform,figPowerPsiExp}).
The \modelDPP and \modelUniform models estimate low posterior probability of
$\divTimeNum = 1$ across all of the simulations.
This is in contrast to the \modelOld model, which infers high posterior probabilities of a single
divergence for most replicates across all simulations
(Figure~\ref{figPowerPsiProbOld4} \& 
S\labelcref{figPowerPsiProbOld,figPowerPsiUniform,figPowerPsiExp}).
The exponential priors on divergence-time and theta parameters (model
\modelUshaped) result in lower estimates of the probability of one divergence
when compared to \modelOld, but higher estimates when compared to \modelUniform
and \modelDPP
(Figure~\ref{figPowerPsiProbOld4} \& 
S\labelcref{figPowerPsiProbOld,figPowerPsiUniform,figPowerPsiExp}).
The \modelDPP and \modelUniform models do frequently support the one-model
according to a Bayes factor criterion of greater than 10, however still less
frequently than the \modelOld model.
This result is not surprising given the extremely small prior probability of
the one-divergence model under the \modelDPP and \modelUniform models (I.e.,
very few posterior samples of the one-divergence model will result in a
large Bayes factor under these models).
However, the small posterior probability of the one-divergence model estimated
under \modelDPP and \modelUniform should prevent an investigator from
overinterpreting the Bayes factor as strong support for clustered divergences.

Lastly, when we look at the estimated posterior probability of
\divTimeDispersion being consistent with one shared divergence
($p(\divTimeDispersion < 0.01 | \ssSpace)$), we see the same pattern of model
behavior, with \modelDPP and \modelUniform inferring low probabilities across
all simulations, \modelOld inferring high probabilities, and \modelUshaped
inferring intermediate values
(Figure~\ref{figPowerOmegaProbOld4} \& 
S\labelcref{figPowerOmegaProbOld,figPowerOmegaUniform,figPowerOmegaExp}).

@@

\section*{Discussion}
Our results demonstrate that using alternative priors on parameters and
divergence models improved the behavior of the \msb model.
In our implementation, model-choice estimation is more accurate and shows
greater robustness to model violations
(Figure~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).
The original model is very sensitive to model violations and, when present,
strongly over-estimates the probability of one divergence event shared
across all taxa
(Figure~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).
When more appropriate priors are used for divergence-time and theta parameters,
and either a Dirichlet-process or uniform prior applied across divergence
models, the model is less sensitive to violations, and, when they do cause
bias, the method is conservative and tends to under-estimate the probability of
biogeographically interesting models of temporally clustered divergences
(Figure~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).

Furthermore, the modifications to the model also improve the method's power to
detect random variation in divergence times, reducing the bias toward
estimating clustered divergences
(Figures~\labelcref{figPowerPsiOld4,figPowerPsiProbOld4,figPowerOmegaOld4,figPowerOmegaProbOld4}.
Our results are similar to those of \citet{Oaks2012} in that we also find
the \msb model to infer strong support for clustered divergences when
divergences are random over quite broad timescales
(Figures~\labelcref{figPowerPsiOld4,figPowerPsiProbOld4,figPowerOmegaOld4,figPowerOmegaProbOld4}.
Our results expand on this by showing that this behavior is consistent
across a range of conditions underlying the data.
When using the new priors, the method is less prone to spurious inference of
clustered divergence models, and much less likely to incorrectly infer such
models with strong support.

By evaluating a model intermediate between the old and new implementation
(\modelUshaped), we were able to determine the relative effects of the
modifications to the model.
Across all of our analyses, the results show that using better priors on
divergence-time and nuisance parameters alone does improve the performance of
the method.
The magnitude of the bias toward inferring support for the one-divergence
model when there are model violations is reduced when the exponential
priors are used in place of the uniform priors
(Figure~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).
Furthermore, better priors on nuisance parameters improves the method's
power to detect temporally random divergences.
Throughout our analyses, the intermediate model (\modelUshaped) performs
better than the \msb model, but not as well as the models with alternative
priors on divergence models.
This suggests, as predicted by \citet{Oaks2012}, that the bias of \msb toward
models of temporally clustered divergences is caused by a combination of
(1) uniform priors on parameters hindering the marginal likelihoods of models
with more \divTime{} parameters and
(2) the U-shaped prior on divergence models giving low prior density to models
with intermediate numbers of divergence parameters.

While our modifications improve the behavior of the model, we urge
caution when using the method and interpreting its results.
The method attempts to approximate the posterior of a very parameter-rich model
using relatively little information from the data.
For example, when applied to the dataset of Philippine vertebrates analyzed by
\citet{Oaks2012} the model can have as many as 604--625 free parameters
(depending on \divTimeNum), and samples over 1002 divergence models.
Our modifications allow this model to be simplified to 471--492 free
parameters.
Sampling over hundreds of parameters and more than a thousand models is a
difficult inference problem, especially when only using a small number of
summary statistics calculated from the sequence alignments of each taxon pair.
The population-genetic summary statistics used by our method contain little
information about many of the free parameters in the model.
Thus, we expect the improved method will still be sensitive to priors, and the
power, while improved, may still be low.
Furthermore, ABC methods in general have been shown to be biased
\citep{Robert2011} for model choice.
As recommended by \citet{Oaks2012}, any results from the method should be
accompanied by assessments of prior sensitivity and power.
We encourage investigators to view this method and its predecessor as a means
of exploring their data rather than a rigorous means of evaluating hypotheses.

### 2013-12-19 manuscripts/dpp-msbayes/dpp-msbayes-tables.tex
Working on DPP ms.
Finishing incorporation of the new figures. Also, updating some of the
new figures to add an extra digit for the exponential distributions of
divergence times.

    % \footnotesize

@@

        assessing the power of the models in Table \ref{tabPriors}.}

@@

    \begin{tabular}{ l l l l l }

@@

        & \multicolumn{4}{c}{Priors} \\
        \cmidrule(){2-5}
        Model series & \divTimeIndexVector & \divTime{} & \ancestralTheta{} & \descendantTheta{}{} \\

@@

@@

@@

@@

@@

                            & $\divTime{} \sim Exp(17.3205)$

@@

                            & $\divTime{} \sim Exp(8.6603)$

@@

                            & $\divTime{} \sim Exp(5.7735)$

@@

                            & $\divTime{} \sim Exp(4.3301)$

@@

                            & $\divTime{} \sim Exp(3.4641)$

@@

                            & $\divTime{} \sim Exp(1.7321)$

>>>>

    \footnotesize

@@

        assessing the power of the models in Table \ref{tabPriors}.
        Distributions of divergence times are given in units of \globalcoalunit
        generations and millions of generations ago (MGA), the latter
        assuming a per-site rate of 1\e{-8} mutations per generation.}

@@

    \begin{tabular}{ l l l l l l }

@@

        & \multicolumn{5}{c}{Priors} \\
        \cmidrule(){2-6}
        Model series & \divTimeIndexVector & \divTime{} (\globalcoalunit gens.) & \divTime{} (MGA) & \ancestralTheta{} & \descendantTheta{}{} \\

@@

                            & $\divTime{} \sim U(0,0.5)$

@@

                            & $\divTime{} \sim U(0,1.5)$

@@

                            & $\divTime{} \sim U(0,2.5)$

@@

                            & $\divTime{} \sim U(0,5.0)$

@@

                            & $\divTime{} \sim Exp(17.32)$
                            & $\divTime{} \sim Exp(mean = 0.14)$

@@

                            & $\divTime{} \sim Exp(8.66)$
                            & $\divTime{} \sim Exp(mean = 0.29)$

@@

                            & $\divTime{} \sim Exp(5.77)$
                            & $\divTime{} \sim Exp(mean = 0.43)$

@@

                            & $\divTime{} \sim Exp(4.33)$
                            & $\divTime{} \sim Exp(mean = 0.58)$

@@

                            & $\divTime{} \sim Exp(3.46)$
                            & $\divTime{} \sim Exp(mean = 0.72)$

@@

                            & $\divTime{} \sim Exp(1.73)$
                            & $\divTime{} \sim Exp(mean = 1.44)$
